---
title: "HW4_Salem"
author: "Mohamed Salem"
date: "September 24, 2019"
output: pdf_document
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=60),tidy=TRUE,fig.height = 4.4, fig.width = 6, fig.align = 'center')
```

```{r , echo=F,results='hide', collapse=TRUE, include=FALSE}
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(formatR))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lintr))
```

I concluded that there is value in adopting a specific style. Personally, given my current coding style, I've resolved to evolve my style to follow the conventions of using BigCamelCase and not using the attach function.


```{r , echo=T, collapse=TRUE, include=FALSE}
require(lintr)
lint(filename = "D:/Vtech/Statistical Programming/HW3_Salem.Rmd")
```

The main recommendation I received was to add spaces between my operators. I was also pointed towards not having my code exceed 80 characters per line.

```{r , echo=T,tidy.opts=list(width.cutoff=65),tidy=TRUE}

#Read and import the object
data<-readRDS(file = "D:/Vtech/Statistical Programming/HW4_data.rds")

#A function that takes the dataset and performs a number of prespecified exploratory analyses on them
AnalyzeData<-function(x) {
  SumData<-x
  names(SumData)<-c("Observer","Var_1","Var_2")
  SumData<-SumData %>%
    group_by(Observer)%>%
    summarise(
      dev1_mean=mean(Var_1,na.rm = T),
      dev2_mean=mean(Var_2,na.rm = T),
      dev1_sd=sd(Var_1,na.rm = T),
      dev2_sd=sd(Var_2,na.rm = T),
      dev_corr=cor(Var_1,Var_2, method = "pearson")
    ) 
  
  SumDataMelted<-data.frame(matrix(rep(c(1,2), each = length(SumData$dev1_mean),nrow = length(SumData$dev1_mean), ncol=1)))
  names(SumDataMelted)<-c("Variable")
  RowMeltmean<-rbind(data.frame(means=SumData$dev1_mean),data.frame(means=SumData$dev2_mean))
  RowMeltsd<-rbind(data.frame(means=SumData$dev1_sd),data.frame(means=SumData$dev2_sd))
  SumDataMelted<-data.frame(cbind(as.factor(SumDataMelted$Variable),RowMeltmean$means,RowMeltsd))
  names(SumDataMelted)<-c("Variable","MeanValues","SdValues")
  SumDataMelted$Variable<-as.factor(SumDataMelted$Variable)
  
  bxplt <- ggplot(data = SumDataMelted, aes(x=Variable, y=MeanValues)) +
                geom_boxplot()
  
  plotB <- bxplt + facet_wrap( ~ Variable, scales="free")
  
  vlnplot <- ggplot(data = SumDataMelted, aes(x=Variable, y=SdValues)) + 
                geom_violin(trim = FALSE)
  
  plotA <- vlnplot + facet_wrap( ~ Variable, scales="free")
  
  return(list(SumData, plotB, plotA))
}
```
\newpage
```{r echo=TRUE,tidy.opts=list(width.cutoff=65),tidy=TRUE}
#Testing out our function
AnalyzeData(data)
```
\newpage
```{r echo=TRUE,tidy.opts=list(width.cutoff=65),tidy=TRUE}
#A function to implement the Riemann sum method of finding areas with the arguments a, b, f, eps, N, representing the starting point of a continuous domain, the ending point of a continuous domain, our function, our tolerance level, and maximum number of iterations, respectively. 
RamenSum <- function(a,b,f,eps=1E-6, N=10000) {
  i = 1
  h = 1
  SliceWidths = numeric(N)
  SliceWidths[1] = h
  IntrmSoln = numeric(N)
  RSum = 0
  cdf = integrate(f, 0, 1)
  while (abs(RSum - cdf$value) > eps) {
    Pts = seq(a,b, by = h)
    NumPts = length(Pts)
    midpts = 0.5*(Pts[2:NumPts]+Pts[1:(NumPts-1)])
    RSum = sum(h*f(midpts))
    h = 0.5*h
    i=i+1
    SliceWidths[i] = h
  }
  return(list(paste("Solution = ", RSum), paste("Soln Slice Width = ",h), paste("True Integral = ",cdf$value),SliceWidths[1:(i-1)]))
}
```

```{r echo=TRUE,tidy.opts=list(width.cutoff=65),tidy=TRUE}
#Testing out our function
f <- function(x) {exp(-(x^2)/2)}
RamenSum(0,1,f)
```
\newpage
```{r echo=TRUE,tidy.opts=list(width.cutoff=65),tidy=TRUE}
#A function to implement the Newton-Rhapson method of finding roots with the arguments f, eps, x0, N, representing our function, our tolerance level, our initial guess, and maximum number of iterations, respectively. Function adapted from code by Yin Zhao, available at: https://www.academia.edu/7031789/Newton-Raphson_Method_in_R
#Adapted on 9/24/2019

NewtonRhapson <- function(f, eps=1E-3, x0 = 1, N=10000) {
  h = 1e-3
  i = 1
  x1 = x0
  xstatic = x0
  p = numeric(N)
  while (i<=N) {
    derv = (f(x0 + h)-f(x0))/h
    x1 = (x0 - (f(x0)/derv))
    p[i] = x1
    i = i+1
    if (abs(x1 - x0) < eps) break
    x0 = x1
  }
  
  m<-seq(-50,50,length=10000)
  RootGrph<-plot(p, f(p), type='l', lwd=1.5, main=expression(f(Newton-Rhapson),xlim = c(-9,9), ylim = c(-5,5), ylab = "f(x)"))
  abline(h=0)
  FnGrph<-plot(m, f(m), type='l', lwd=1.5, main=expression(f(Domain)),xlim = c(-9,9), ylim = c(-5,5), ylab = "f(x)")
  abline(h=0)
  
  return(list(paste("Solution = ", x1), paste("tolerance = ",eps),paste("Initial Guess = ",xstatic),paste("Initial Interval = ",xstatic, " to ",xstatic+h),Solution_Iterations = p[1:(i-1)],RootGrph, FnGrph))
}
```

```{r echo=TRUE,tidy.opts=list(width.cutoff=65),tidy=TRUE}
#Testing Out Our Function
f <- function(x) {3^x - sin(x) + cos(5*x)}
NewtonRhapson(f)
```